---
title: "A Latent Profile Analysis of Attachment Dimensions and Their Associations with Sleep, Mood, and Fatigue"
author: Iris Chen
format:
  html:
    embed-resources: true
    code-fold: true
    code-tools: true
    toc: true
execute:
  df-print: kable
  message: false
  warning: false
  cache: false
---

## Data Management

```{r}
# load up all the packages we need for this analysis
library(tidyLPA)
library(descr)
library(dplyr)
library(tidyr)
library(corrplot)
library(qacBase)
library(Hmisc)
library(gtsummary)
library(apaTables)
library(psych)
library(ggplot2)

# read in the raw survey data
data = read.csv("SleepHygiene.csv")
```

```{r}
# first row is just the question text from the survey, not real answers, so drop it
data <- data[-1, ]
rownames(data) <- NULL
```

### Rename Columns

```{r}
# the original column names (Q1, Q18_1REC, etc.) are basically unreadable,
# so renaming everything to something that actually makes sense
data <- data %>%
  rename(
    gender = Q1,
    relationship_duration = Q2_A,
    age = Q3_A,
    ethnicity = Q4_A,
    live = Q32,
    hygiene_1 = Q18_1REC,
    hygiene_2 = Q18_2REC,
    hygiene_3 = Q18_3REC,
    hygiene_4 = Q18_4REC,
    hygiene_5 = Q18_5REC,
    hygiene_6 = Q18_6rec,
    hygiene_7 = Q18_7rec,
    hygiene_8 = Q18_8rec,
    hygiene_9 = Q18_9rec,
    hygiene_10 = Q18_10rec,
    hygiene_11 = Q18_11rec,
    hygiene_12 = Q18_12rec,
    hygiene_13 = Q18_13rec,
    SLEEP_HYG_SCORE = Health_Behaviour,
    fatigue = Q20_1,
    sleep_duration = leep_Duration,
    bedtime = Q27,
    wakeuptime = Q28,
    insomnia_psqi = Q29_1_RECODE,
    wakeful_psqi = Q29_2_REC,
    breathing_psqi = Q29_3_RECODE,
    snoring_psqi = Q29_4_RECODE,
    hot_psqi = Q29_5_RECODE,
    nightmares_psqi = Q29_6_RECODE,
    pain_psqi = Q29_7_RECODE,
    Anxious1 = Anxiety1
  )
```

### Select Columns

```{r}
# just grabbing the columns we actually need for this analysis, ignoring the rest
library(dplyr)
data_select <- data %>%
  select(gender, relationship_duration, age, ethnicity, live,
         fatigue, bedtime, wakeuptime, sleep_duration, SleepEfficiency,
         SLEEP_DUR_SCORE, ONSET_SCORE, SLP_QUALITY_SCORE,
         pain_psqi, nightmares_psqi, hot_psqi, snoring_psqi, breathing_psqi,
         wakeful_psqi, insomnia_psqi, Sleep_Problem_total, Sleep_Problem_score,
         Enthusiasm2, TroubleAWAKE, ENTHUSIASMTROUBLE,
         Enthusiasm_and_trouble_score, sleep_efficiencySCORE, SLEEP_SCORE_TOTAL,
         Avoidance1REC, Avoidance3, Avoidance5REC, Avoidance7, Avoidance9REC,
         Avoidance11, Anxiety2, Anxiety4, Anxiety6, Anxiety8REC, Anxiety10,
         Anxiety12,
         AnxietyMEAN, AvoidanceMEAN, Avoidance, Anxiety, hygiene_1, hygiene_2,
         hygiene_3, hygiene_4, hygiene_5, hygiene_6, hygiene_7, hygiene_8,
         hygiene_9, hygiene_10, hygiene_11, hygiene_12, hygiene_13, SLEEP_HYG_SCORE,
         Anxious1, Anxious2, depressed1, depressed2, Anxiousmood, Depressedmood)
```

### Convert Null to NA

```{r}
# the missing values from the survey export show up as the weird string "#NULL!",
# R doesn't recognize that as missing, so swap it out for a real NA
data_select[data_select == "#NULL!"] <- NA
```

### Convert Datatypes

```{r}
# these ones are categorical, so turn them into factors
data_select <- data_select %>%
  mutate(
    gender = factor(gender),
    relationship_duration = factor(relationship_duration),
    ethnicity = factor(ethnicity),
    live = factor(live)
  )
```

```{r}
# everything except those factor columns and the two time columns should be numeric
data_select <- data_select %>%
  mutate(across(-c(bedtime,
                   wakeuptime,
                   gender,
                   relationship_duration,
                   ethnicity,
                   live),
                as.numeric))
```

## Correlation Analysis

```{r}
# grabbing a bunch of sleep / mood / attachment variables to see how they correlate with each other
cor_data <- data_select %>%
  select(relationship_duration,
         SLEEP_DUR_SCORE,
         ONSET_SCORE,
         SLP_QUALITY_SCORE,
         Sleep_Problem_score,
         Enthusiasm_and_trouble_score,
         sleep_efficiencySCORE,
         Anxious1,
         Anxious2,
         depressed1,
         depressed2,
         Avoidance1REC,
         Avoidance3,
         Avoidance5REC,
         Avoidance7,
         Avoidance9REC,
         Avoidance11,
         Anxiety2,
         Anxiety4,
         Anxiety6,
         Anxiety8REC,
         Anxiety10,
         Anxiety12,
        Sleep_Problem_total,
        SLEEP_SCORE_TOTAL,
        Avoidance,
        Anxiety,
        SLEEP_HYG_SCORE,
        Anxiousmood,
        Depressedmood)

# quick correlation plot so we can see what's related to what at a glance
cor_plot(cor_data)
```

## Descriptive Statistics

```{r}
# cutting age into groups so we can do demographic descriptives
data_des <- data_select %>%
  mutate(
    age_group = cut(
      age,
      breaks = c(17,29,39,49,59,69,79,89),
      labels = c("18-29","30–39","40–49","50–59",
                 "60–69","70–79","80–89"),
      right = TRUE
    )
  )

# basic demographics table (age group / gender / ethnicity / relationship duration / living situation)
table1 <- data_des %>%
  select(age_group, gender, ethnicity, relationship_duration, live) %>%
  tbl_summary(
    statistic = all_categorical() ~ "{p}%",
    missing = "no"
  ) %>%
  modify_header(label = "**Category**") %>%
  bold_labels()

table1
```

```{r}
# pulling out the continuous variables separately to run descriptives + a correlation table, then exporting to Table2
con_vars <- data_select %>%
  select(sleep_duration,
         SleepEfficiency,
         SLEEP_DUR_SCORE,
         Anxiety,
         Avoidance,
         SLEEP_HYG_SCORE,
         Anxiousmood,
         Depressedmood,
         fatigue)

cor_results <- corr.test(con_vars, use = "pairwise")
desc <- psych::describe(con_vars)[, c("mean", "sd")]

# export an APA-style correlation table (Table2.doc)
# apa.cor.table(con_vars,
#               filename = "Table2.doc",
#               show.conf.interval = FALSE)
```

## LPA Analysis

```{r}
# giving the attachment scale items friendlier names - AV = avoidance items, AX = anxiety items
lpa_rename <- data_select %>%
  rename(
    AV1_TurnNeed = Avoidance1REC,
    AV3_PullBack = Avoidance3,
    AV5_Comfort = Avoidance5REC,
    AV7_AvoidClose = Avoidance7,
    AV9_Discuss = Avoidance9REC,
    AV11_NervousClose = Avoidance11,
    AX2_Reassure = Anxiety2,
    AX4_NotClose = Anxiety4,
    AX6_ScareAway = Anxiety6,
    AX8_NoAbandon = Anxiety8REC,
    AX10_Frustrated = Anxiety10,
    AX12_WorryCare = Anxiety12
    )
```

```{r}
# run 1 to 4 classes first to see which number of profiles fits best (compare AIC/BIC/entropy etc.)
set.seed(1234)
vLPA <- lpa_rename %>%
  select(AV1_TurnNeed,
         AV3_PullBack,
         AV5_Comfort,
         AV7_AvoidClose,
         AV9_Discuss,
         AV11_NervousClose,
         AX2_Reassure,
         AX4_NotClose,
         AX6_ScareAway,
         AX8_NoAbandon,
         AX10_Frustrated,
         AX12_WorryCare) %>%
  single_imputation() %>%
  scale()  %>%
  estimate_profiles(1:4, nrep = 50)

vLPA
plot_profiles(vLPA, solution = 4)

# quick summary of the fit stats so we can compare across models
fit <- get_fit(vLPA)
fit %>%
  dplyr::select(Model, Classes, AIC, BIC, Entropy, n_min, n_max, BLRT_val, BLRT_p)
```

```{r}
# going with the 4-class solution, rerunning it on its own
set.seed(1234)
Class4LPA <- lpa_rename %>%
  select(AV1_TurnNeed,
         AV3_PullBack,
         AV5_Comfort,
         AV7_AvoidClose,
         AV9_Discuss,
         AV11_NervousClose,
         AX2_Reassure,
         AX4_NotClose,
         AX6_ScareAway,
         AX8_NoAbandon,
         AX10_Frustrated,
         AX12_WorryCare) %>%
  single_imputation() %>%
  scale()  %>%
  estimate_profiles(4)

Class4LPA
plot_profiles(Class4LPA)
```

```{r}
# pull out each person's assigned class along with the original data
plot_data <- get_data(Class4LPA)
plot_data
```

### Posterior Probability

```{r}
# find the column names that hold the posterior probabilities, need those next
prob_cols <- names(plot_data)[grepl("prob|cprob|posterior", names(plot_data), ignore.case = TRUE)]

prob_cols
```

```{r}
# peek at the first 10 people's probabilities of belonging to each class
plot_data %>%
  dplyr::select(Class, all_of(prob_cols)) %>%
  head(10)
```

```{r}
# average posterior probability by class - checking how "clean" the classification actually is
pp_by_class <- plot_data %>%
  group_by(Class) %>%
  summarise(
    across(all_of(prob_cols), ~ mean(.x, na.rm = TRUE)),
    n = n(),
    .groups = "drop"
  )

pp_by_class
```

```{r}
# for each person, grab the highest posterior probability (the one for the
# class they were actually assigned to) - closer to 1 = more confident classification
plot_data <- plot_data %>%
  mutate(
    max_postprob = do.call(pmax, c(across(all_of(prob_cols)), na.rm = TRUE))
  )

# summarize that max probability by class (mean/sd/min/max)
plot_data %>%
  group_by(Class) %>%
  summarise(
    n = n(),
    mean_max_postprob = mean(max_postprob, na.rm = TRUE),
    sd_max_postprob = sd(max_postprob, na.rm = TRUE),
    min_max_postprob = min(max_postprob, na.rm = TRUE),
    max_max_postprob = max(max_postprob, na.rm = TRUE),
    .groups = "drop"
  )
```

```{r}
# grab the data for the 4-class solution
vLPA4 <- get_data(vLPA, solution = 4)

# reshape to long format so it's easier to plot (one row per item)
long_data <- vLPA4 %>%
  pivot_longer(
    cols = starts_with("AV") | starts_with("AX"),
    names_to = "Indicator",
    values_to = "Score"
  )
long_data
```

### Calculate Means per Class

```{r}
# average score per class per item, need this for the profile plots
plot_lpa <- long_data %>%
  group_by(Class, Indicator) %>%
  summarise(Mean = mean(Score), .groups = "drop")
```

```{r}
# lock in the order the items show up in on the plot, otherwise it defaults
# to alphabetical and looks messy
plot_lpa <- plot_lpa %>%
  mutate(
    Indicator = factor(
      Indicator,
      levels = c(
        "AV1_TurnNeed",
        "AV3_PullBack",
        "AV5_Comfort",
        "AV7_AvoidClose",
        "AV9_Discuss",
        "AV11_NervousClose",
        "AX2_Reassure",
        "AX4_NotClose",
        "AX6_ScareAway",
        "AX8_NoAbandon",
        "AX10_Frustrated",
        "AX12_WorryCare"
      )
    )
  )
```

```{r}
# use line type to tell the four profiles apart, plotting the standardized
# mean score for each item
ggplot(plot_lpa, aes(x = Indicator, y = Mean,
                      group = Class,
                      linetype = factor(Class))) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  scale_linetype_manual(
  values = c("solid", "dashed", "dotted", "dotdash"),
  labels = c("Dismissive-leaning", "Fearful-avoidant", "Anxious-preoccupied", "Secure")
) +
  theme_classic() +
  labs(
    y = "Standardized Scores",
    x = "Attachment Indicators",
    linetype = "Profile"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
# same plot but with color instead of linetype - a bit easier to read
plot_lpa$Class <- factor(plot_lpa$Class)

ggplot(plot_lpa,
       aes(x = Indicator,
           y = Mean,
           group = Class,
           color = Class)) +

  geom_line(size = 1) +
  geom_point(size = 2) +

  scale_color_manual(values = c(
    "#1b9e77",
    "#d95f02",
    "#7570b3",
    "#e7298a",
    "#66a61e",
    "#e6ab02"
  )) +

  theme_classic() +

  labs(
    y = "Standardized Scores",
    x = "Attachment Indicators",
    color = "Profile"
  ) +

  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
# faceted version - splits the four profiles into separate small plots so
# it's easier to look at each one
plot_lpa$Class <- factor(plot_lpa$Class)

ggplot(plot_lpa,
       aes(x = Indicator,
           y = Mean,
           group = Class,
           color = Class)) +

  geom_line(size = 1) +
  geom_point(size = 2) +
  geom_hline(aes(yintercept = 0), linetype = 2)+
  facet_wrap(~Class)+

  scale_color_manual(values = c(
    "#1b9e77",
    "#d95f02",
    "#7570b3",
    "#e7298a",
    "#66a61e",
    "#e6ab02"
  )) +

  theme_classic() +

  labs(
    y = "Standardized Scores",
    x = "Attachment Indicators",
    color = "Profile"
  ) +

  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### Profile Interpretation

Based on the plots above, here's the plain-language read on the four profiles (not a formal definition, just what the pattern seems to say):

**Profile 1 (low anxiety, moderate avoidance)**

- Turns to partner less
- Seeks comfort less
- Shares problems less

These people aren't really anxious about the relationship, but they also don't lean on their partner much for emotional support - kind of independent, a bit distant.

**Profile 2 (fearful/disorganized attachment)**

A few items are reverse-coded (Q1, Q5, Q9), so they look low on the graph, but once you flip them back:

- Turns to partner a lot (Q1)
- Really wants comfort/reassurance (Q5)
- Very willing to talk about problems (Q9)

So underneath, there's actually a strong desire for closeness. At the same time:

- Very high anxiety: needs reassurance (Q2), worries about partner's feelings (Q12), gets frustrated when partner's unavailable (Q10), feels rejected (Q4)
- Very high avoidance: pulls back (Q3), avoids closeness (Q7), gets nervous when partner gets close (Q11), scares partners away (Q6)

Basically: really wants closeness and reassurance, but also feels unsafe and pulls away - classic fearful/disorganized attachment.

**Profile 3 (anxious/preoccupied attachment)**

- Low avoidance: doesn't avoid closeness (Q7), not very nervous about closeness (Q11), doesn't pull back much (Q3)
- High anxiety: needs reassurance (Q2), worries partner doesn't care enough (Q12), gets frustrated when unavailable (Q10), feels not close enough (Q4)
- Lots of closeness behaviors: turns to partner (Q1, after reversing), seeks comfort (Q5, after reversing), talks about problems (Q9, after reversing)

Short version: wants closeness and doesn't avoid it, but worries a lot about the relationship.

**Profile 4 (secure)**

Both anxiety and avoidance are low - the most "stable" group of the four.

## ANOVA & post hoc

```{r}
# check how many people ended up in each class
plot_data$Class <- factor(plot_data$Class)
table(plot_data$Class)
```

```{r}
# save the 12 attachment item names, need this for the loop below
vars <- c(
  "AV1_TurnNeed",
  "AV3_PullBack",
  "AV5_Comfort",
  "AV7_AvoidClose",
  "AV9_Discuss",
  "AV11_NervousClose",
  "AX2_Reassure",
  "AX4_NotClose",
  "AX6_ScareAway",
  "AX8_NoAbandon",
  "AX10_Frustrated",
  "AX12_WorryCare"
)
```

```{r}
# average score per item per class, using the standardized data (same numbers as what's in the plots)
mean_table <- plot_data %>%
  group_by(Class) %>%
  summarise(across(all_of(vars), ~ mean(.x, na.rm = TRUE)))

mean_table
```

```{r}
# switching to the raw (non-standardized) scores here, easier to interpret for the statistical tests
raw_data_with_class <- lpa_rename %>%
  select(AV1_TurnNeed,
         AV3_PullBack,
         AV5_Comfort,
         AV7_AvoidClose,
         AV9_Discuss,
         AV11_NervousClose,
         AX2_Reassure,
         AX4_NotClose,
         AX6_ScareAway,
         AX8_NoAbandon,
         AX10_Frustrated,
         AX12_WorryCare) %>%
  single_imputation() %>%
  mutate(Class = plot_data$Class)
```

```{r}
# Class needs to be a factor or the ANOVA below will complain
raw_data_with_class$Class <- factor(raw_data_with_class$Class)
```

```{r}
# raw-score version of the mean table, for comparing against the standardized one
mean_table_raw <- raw_data_with_class %>%
  group_by(Class) %>%
  summarise(across(all_of(vars), mean, na.rm = TRUE))

mean_table_raw
```

```{r}
# testing the workflow on one variable first - one-way ANOVA + post hoc -
# just to make sure everything works before looping
model1 <- aov(AV1_TurnNeed ~ Class, data = raw_data_with_class)
summary(model1)
TukeyHSD(model1)
```

```{r}
# turning that workflow above into a loop, running ANOVA on all 12 attachment items
library(purrr)
anova_results <- purrr::map_dfr(vars, function(v) {
  model <- aov(as.formula(paste(v, "~ Class")), data = raw_data_with_class)
  out <- summary(model)[[1]]

  tibble(
    Indicator = v,
    F_value = out$`F value`[1],
    p_value = out$`Pr(>F)`[1]
  )
})

anova_results
```

```{r}
# using compact letter display to see which classes differ significantly:
# same item, different letters = those two classes are significantly different
library(multcomp)
library(emmeans)
cld_results <- purrr::map_df(vars, function(v) {
  model <- aov(as.formula(paste(v, "~ Class")), data = raw_data_with_class)
  emm <- emmeans(model, ~ Class)
  out <- cld(emm, adjust = "tukey", Letters = letters) %>%
    as.data.frame()

  out %>%
    transmute(
      Indicator = v,
      Class,
      Mean = emmean,
      Group = gsub(" ", "", .group)
    )
})

cld_results
```

```{r}
# reshaping the item x class means and significance letters into a wide table,
# so it's ready to drop straight into the paper
means_wide <- cld_results %>%
  dplyr::select(Indicator, Class, Mean) %>%
  pivot_wider(
    names_from = Class,
    values_from = Mean
  ) %>%
  rename(
    Profile1_Mean = `1`,
    Profile2_Mean = `2`,
    Profile3_Mean = `3`,
    Profile4_Mean = `4`
  )

groups_wide <- cld_results %>%
  dplyr::select(Indicator, Class, Group) %>%
  pivot_wider(
    names_from = Class,
    values_from = Group
  ) %>%
  rename(
    Profile1_Group = `1`,
    Profile2_Group = `2`,
    Profile3_Group = `3`,
    Profile4_Group = `4`
  )

# also building a "high to low" significance ordering, like "2 > 3 > 1,4"
ordering_table <- cld_results %>%
  arrange(Indicator, desc(Mean)) %>%
  group_by(Indicator, Group) %>%
  summarise(
    Classes = paste(Class, collapse = ","),
    Mean = mean(Mean),
    .groups = "drop"
  ) %>%
  arrange(Indicator, desc(Mean)) %>%
  group_by(Indicator) %>%
  summarise(
    Sig_Order = paste(Classes, collapse = " > "),
    .groups = "drop"
  )

final_table <- anova_results %>%
  left_join(means_wide, by = "Indicator") %>%
  left_join(groups_wide, by = "Indicator") %>%
  left_join(ordering_table, by = "Indicator")

final_table

# could export to csv if needed, leaving it commented out for now
# write.csv(final_table, "Table5_style_attachment_results.csv", row.names = FALSE)
```

```{r}
# paper-style formatted table, exported to a Word doc (Table5)
library(officer)
library(flextable)

table_for_doc <- final_table %>%
  mutate(
    Profile1_Mean = round(Profile1_Mean, 2),
    Profile2_Mean = round(Profile2_Mean, 2),
    Profile3_Mean = round(Profile3_Mean, 2),
    Profile4_Mean = round(Profile4_Mean, 2),
    F_value = round(F_value, 2),
    p_value = ifelse(as.numeric(p_value) < .001, "< .001", sprintf("%.3f", as.numeric(p_value)))
  ) %>%
  rename(
    Dimension = Indicator,
    `Profile 1` = Profile1_Mean,
    `Profile 2` = Profile2_Mean,
    `Profile 3` = Profile3_Mean,
    `Profile 4` = Profile4_Mean,
    `F` = F_value,
    `p` = p_value,
    `Sig. Order` = Sig_Order
  ) %>%
  dplyr::select(Dimension, `Profile 1`, `Profile 2`, `Profile 3`, `Profile 4`, `F`, `p`, `Sig. Order`)

ft <- flextable(table_for_doc) %>%
  theme_booktabs() %>%
  autofit() %>%
  align(j = 1, align = "left", part = "all") %>%
  align(j = 2:ncol(table_for_doc), align = "center", part = "all") %>%
  bold(part = "header") %>%
  fontsize(size = 11, part = "all") %>%
  font(fontname = "Times New Roman", part = "all") %>%
  padding(padding = 4, part = "all")

doc <- read_docx() %>%
  body_add_par("Table 5", style = "Normal") %>%
  body_add_par("Raw mean attachment dimension scores for the four-profile solution.", style = "Normal") %>%
  body_add_flextable(ft) %>%
  body_add_par(
    "Note. Values are raw means. Sig. Order indicates significant pairwise differences across profiles based on Tukey-adjusted comparisons.",
    style = "Normal"
  )

print(doc, target = "Table5_attachment_results.docx")
```

## Post Analysis

```{r}
# giving everyone an id so we can merge the Class assignment back into the main dataset later
data_select <- data_select %>%
  mutate(id = row_number())

lpa_input <- data_select %>%
  rename(
    AV1_TurnNeed = Avoidance1REC,
    AV3_PullBack = Avoidance3,
    AV5_Comfort = Avoidance5REC,
    AV7_AvoidClose = Avoidance7,
    AV9_Discuss = Avoidance9REC,
    AV11_NervousClose = Avoidance11,
    AX2_Reassure = Anxiety2,
    AX4_NotClose = Anxiety4,
    AX6_ScareAway = Anxiety6,
    AX8_NoAbandon = Anxiety8REC,
    AX10_Frustrated = Anxiety10,
    AX12_WorryCare = Anxiety12
  )

# rerunning the 4-class LPA (same setup as before, so results stay consistent)
set.seed(1234)
Class4LPA <- lpa_input %>%
  dplyr::select(AV1_TurnNeed, AV3_PullBack, AV5_Comfort, AV7_AvoidClose,
         AV9_Discuss, AV11_NervousClose, AX2_Reassure, AX4_NotClose,
         AX6_ScareAway, AX8_NoAbandon, AX10_Frustrated, AX12_WorryCare) %>%
  single_imputation() %>%
  scale() %>%
  estimate_profiles(4)

# grab everyone's class assignment and merge it back in by id
class_df <- get_data(Class4LPA) %>%
  mutate(id = data_select$id) %>%
  dplyr::select(id, Class)

# now every person's original variables + attachment Class are in one table -
# this is what we'll use for the rest of the analysis
analysis_data <- data_select %>%
  left_join(class_df, by = "id") %>%
  mutate(Class = factor(Class))
```

### Graphs - 2

#### Continuous variables

```{r}
# listing out the continuous variables we want to look at
cont_vars <- c(
  "age",
  "sleep_duration",
  "SleepEfficiency",
  "SLEEP_DUR_SCORE",
  "ONSET_SCORE",
  "SLP_QUALITY_SCORE",
  "Sleep_Problem_score",
  "Enthusiasm_and_trouble_score",
  "sleep_efficiencySCORE",
  "SLEEP_SCORE_TOTAL",
  "SLEEP_HYG_SCORE",
  "Anxiousmood",
  "Depressedmood",
  "fatigue"
)

plot_cont <- analysis_data %>%
  dplyr::select(Class, all_of(cont_vars)) %>%
  pivot_longer(
    cols = -Class,
    names_to = "Variable",
    values_to = "Value"
  )

# violin + boxplot + jittered points all in one, so we get the shape, the
# quartiles, and each individual's value all at once
ggplot(plot_cont, aes(x = Class, y = Value, fill = Class)) +
  geom_violin(trim = FALSE, alpha = .35, na.rm = TRUE) +
  geom_boxplot(width = .15, outlier.shape = NA, alpha = .7, na.rm = TRUE) +
  geom_jitter(width = .08, alpha = .25, size = 1, na.rm = TRUE) +
  facet_wrap(~ Variable, scales = "free_y") +
  theme_classic() +
  labs(
    title = "Continuous variables by attachment profile",
    x = "Attachment profile",
    y = "Value"
  ) +
  theme(
    legend.position = "none",
    strip.text = element_text(size = 9)
  )
```

#### Overall distributions: Histogram

```{r}
# not splitting by class here, just looking at the overall shape of each variable
ggplot(plot_cont, aes(x = Value)) +
  geom_histogram(bins = 25, color = "white", fill = "grey70", na.rm = TRUE) +
  facet_wrap(~ Variable, scales = "free") +
  theme_classic() +
  labs(
    title = "Overall distributions of continuous variables",
    x = "Value",
    y = "Count"
  )
```

```{r}
# now splitting by class to compare distribution shapes across profiles
ggplot(plot_cont, aes(x = Value, fill = Class)) +
  geom_density(alpha = .3, na.rm = TRUE) +
  facet_wrap(~ Variable, scales = "free") +
  theme_classic() +
  labs(
    title = "Density plots by attachment profile",
    x = "Value",
    y = "Density"
  )

# picking a few key variables and using a grid (one row per class) to zoom in
ggplot(filter(plot_cont, Variable %in% c("age","sleep_duration","Sleep_Problem_score")), aes(x = Value, fill = Class)) +
  geom_density(alpha = .3, na.rm = TRUE) +
  facet_grid(Class~Variable, scales = "free") +
  theme_classic() +
  labs(
    title = "Density plots by attachment profile",
    x = "Value",
    y = "Density"
  )
```

#### Categorical demographics: proportion bars

```{r}
# demographic categorical variables - checking their makeup within each attachment profile
cat_vars <- c("gender", "relationship_duration", "ethnicity", "live")

plot_cat <- analysis_data %>%
  dplyr::select(Class, all_of(cat_vars)) %>%
  mutate(across(all_of(cat_vars), as.factor)) %>%
  pivot_longer(
    cols = -Class,
    names_to = "Variable",
    values_to = "Value"
  )

ggplot(plot_cat, aes(x = Class, fill = Value)) +
  geom_bar(position = "fill", na.rm = TRUE) +
  facet_wrap(~ Variable) +
  scale_y_continuous(labels = scales::percent) +
  theme_classic() +
  labs(
    title = "Demographic composition within each attachment profile",
    x = "Attachment profile",
    y = "Proportion",
    fill = "Category"
  )
```

#### Mean pattern plot across variables

```{r}
# standardizing (z-scoring) all the continuous variables so different scales
# can be compared on the same plot
plot_means_df <- analysis_data %>%
  dplyr::select(Class, all_of(cont_vars)) %>%
  mutate(across(all_of(cont_vars), ~ as.numeric(scale(.x)))) %>%
  pivot_longer(
    cols = -Class,
    names_to = "Variable",
    values_to = "Z"
  ) %>%
  group_by(Class, Variable) %>%
  summarise(MeanZ = mean(Z, na.rm = TRUE), .groups = "drop")

# line plot showing the overall pattern per profile across all variables
ggplot(plot_means_df, aes(x = Variable, y = MeanZ, group = Class, color = Class)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  theme_classic() +
  labs(
    title = "Standardized mean patterns across profiles",
    x = NULL,
    y = "Standardized mean (z)"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```

```{r}
# same info, but as a heatmap this time - color depth = z-score, quicker to scan
ggplot(plot_means_df, aes(x = Variable, y = Class, fill = MeanZ)) +
  geom_tile(color = "white") +
  theme_classic() +
  labs(
    title = "Heatmap of standardized means by profile",
    x = NULL,
    y = "Attachment profile",
    fill = "Mean z"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```


#### Missingness

```{r}
# checking how much missing data each continuous variable has
library(scales)
missing_df <- analysis_data %>%
  dplyr::select(all_of(cont_vars)) %>%
  summarise(across(everything(), ~ mean(is.na(.)))) %>%
  pivot_longer(
    cols = everything(),
    names_to = "Variable",
    values_to = "MissingProp"
  )

ggplot(missing_df, aes(x = reorder(Variable, MissingProp), y = MissingProp)) +
  geom_col() +
  coord_flip() +
  scale_y_continuous(labels = percent) +
  theme_classic() +
  labs(
    title = "Missingness by variable",
    x = NULL,
    y = "Percent missing"
  )
```

```{r}
# checking whether missingness depends on attachment Class -
# if a variable is missing a lot more in one class, that could bias the results
library(dplyr)
library(purrr)
library(tidyr)

vars_to_check <- c(
  "age",
  "sleep_duration",
  "SleepEfficiency",
  "SLEEP_DUR_SCORE",
  "ONSET_SCORE",
  "SLP_QUALITY_SCORE",
  "Sleep_Problem_score",
  "Enthusiasm_and_trouble_score",
  "sleep_efficiencySCORE",
  "SLEEP_SCORE_TOTAL",
  "SLEEP_HYG_SCORE",
  "Anxiousmood",
  "Depressedmood",
  "fatigue"
)

missing_by_class <- map_dfr(vars_to_check, \(v) {
  dat <- analysis_data %>%
    mutate(miss = is.na(.data[[v]]))

  tab <- table(dat$Class, dat$miss)

  pval <- tryCatch(chisq.test(tab)$p.value, error = function(e) NA_real_)

  counts <- dat %>%
    group_by(Class) %>%
    summarise(n_nonmissing = sum(!miss), .groups = "drop") %>%
    pivot_wider(names_from = Class, values_from = n_nonmissing, names_prefix = "Class_")

  tibble(
    Variable = v,
    N_used = sum(!dat$miss),
    Missing_p_by_Class = pval
  ) %>%
    bind_cols(counts)
})

missing_by_class
```

- SLEEP_HYG_SCORE (p = .025): missingness for sleep hygiene may not be evenly distributed across attachment profiles.
- For age, SLP_QUALITY_SCORE, and fatigue: the p-values there are basically an artifact of how the test was run, not a real warning sign.

### Explore each variable

#### age

```{r}
ggplot(analysis_data, aes(x = age)) +
  geom_histogram(bins = 15, fill = "grey70", color = "white") +
  facet_wrap(~ Class) +
  theme_classic()

ggplot(analysis_data, aes(x = Class, y = age)) +
  stat_summary(fun = mean, geom = "point", size = 3) +
  stat_summary(fun.data = mean_cl_normal, geom = "errorbar", width = .15) +
  theme_classic()

analysis_data %>%
  group_by(Class) %>%
  summarise(
    n = sum(!is.na(age)),
    mean = mean(age, na.rm = TRUE),
    sd = sd(age, na.rm = TRUE),
    median = median(age, na.rm = TRUE),
    IQR = IQR(age, na.rm = TRUE),
    min = min(age, na.rm = TRUE),
    max = max(age, na.rm = TRUE)
  )
```

```{r}
# Welch ANOVA is mainly about unequal variances (heteroscedasticity) and unequal
# group sizes, and it's usually fairly robust to non-normality, especially with
# moderate sample sizes

oneway.test(age ~ Class, data = analysis_data, var.equal = FALSE)
```

#### gender

```{r}
# checking gender distribution across the four attachment profiles, then
# chi-square test + post hoc
library(chisq.posthoc.test)

table(analysis_data$Class, analysis_data$gender)
prop.table(table(analysis_data$Class, analysis_data$gender), margin = 1)

ggplot(analysis_data, aes(x = Class, fill = gender)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::label_percent()) +
  theme_classic()

tab_gender <- table(analysis_data$Class, analysis_data$gender)
tab_gender

chi_gen <- chisq.test(tab_gender)
chi_gen

Observed_table<-chi_gen$observed

# post hoc test to see exactly which cell (which class x which gender) is
# driving the significance
chisq.posthoc.test(
  Observed_table,
  method = "bonferroni",
  simulate.p.value = TRUE
)

chisq.test(tab_gender)$expected
```

- Profile 1 having more males and fewer females than expected
- Profile 3 having more females and fewer males than expected
- Profiles 2 and 4 do not appear to differ much from expectation

#### ethnicity

```{r}
table(analysis_data$Class, analysis_data$ethnicity)
prop.table(table(analysis_data$Class, analysis_data$ethnicity), margin = 1)

ggplot(analysis_data, aes(x = Class, fill = ethnicity)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::label_percent()) +
  theme_classic()

tab_ethnicity <- table(analysis_data$Class, analysis_data$ethnicity)
chi_eth <- chisq.test(tab_ethnicity)
chi_eth

Observed_table<-chi_eth$observed

chisq.posthoc.test(
  Observed_table,
  method = "bonferroni",
  simulate.p.value = TRUE
)

chisq.test(tab_ethnicity)$expected
```

#### live

```{r}
table(analysis_data$Class, analysis_data$live)
prop.table(table(analysis_data$Class, analysis_data$live), margin = 1)

ggplot(analysis_data, aes(x = Class, fill = live)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::label_percent()) +
  theme_classic()

tab_live <- table(analysis_data$Class, analysis_data$live)
chi_live <- chisq.test(tab_live)
chi_live

Observed_table<-chi_live$observed

chisq.posthoc.test(
  Observed_table,
  method = "bonferroni",
  simulate.p.value = TRUE
)

chisq.test(tab_live)$expected
```

#### relationship duration

```{r}
table(analysis_data$Class, analysis_data$relationship_duration)
prop.table(table(analysis_data$Class, analysis_data$relationship_duration), margin = 1)

ggplot(analysis_data, aes(x = Class, fill = relationship_duration)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::label_percent()) +
  theme_classic()

tab_relationship_duration <- table(analysis_data$Class, analysis_data$relationship_duration)
chi_relationship_duration <- chisq.test(tab_relationship_duration)
chi_relationship_duration

Observed_table<-chi_relationship_duration$observed

chisq.posthoc.test(
  Observed_table,
  method = "bonferroni",
  simulate.p.value = TRUE
)

chisq.test(tab_relationship_duration)$expected
```

> Quick note on where `SLEEP_SCORE_TOTAL` comes from:
> `SLEEP_SCORE_TOTAL = SLEEP_DUR_SCORE + ONSET_SCORE + SLP_QUALITY_SCORE + Sleep_Problem_score + Enthusiasm_and_trouble_score + sleep_efficiencySCORE`

#### Sleep

```{r}
# pulling together the sleep / mood / fatigue scores to compare across attachment profiles
sleep_vars <- c(
  "SLEEP_DUR_SCORE",
  "ONSET_SCORE",
  "SLP_QUALITY_SCORE",
  "Sleep_Problem_score",
  "Enthusiasm_and_trouble_score",
  "sleep_efficiencySCORE",
  "SLEEP_HYG_SCORE",
  "Anxiousmood",
  "Depressedmood",
  "fatigue"
)

sleep_long <- analysis_data %>%
  dplyr::select(Class, all_of(sleep_vars)) %>%
  pivot_longer(
    cols = -Class,
    names_to = "Variable",
    values_to = "Value"
  )

ggplot(sleep_long, aes(x = Class, y = Value, fill = Class)) +
  geom_boxplot(alpha = .75, na.rm = TRUE) +
  facet_wrap(~ Variable, scales = "free_y") +
  theme_classic() +
  labs(
    title = "Sleep, mood, and fatigue scores by attachment profile",
    x = "Attachment profile",
    y = "Score"
  ) +
  theme(
    legend.position = "none",
    strip.text = element_text(size = 9)
  )

# switching to mean + confidence interval to emphasize differences in average level
ggplot(sleep_long, aes(x = Class, y = Value, color = Class)) +
  stat_summary(fun = mean, geom = "point", size = 2.5, na.rm = TRUE) +
  stat_summary(fun.data = mean_cl_normal, geom = "errorbar", width = .15, na.rm = TRUE) +
  facet_wrap(~ Variable, scales = "free_y") +
  theme_classic() +
  labs(
    title = "Mean scores by attachment profile",
    x = "Attachment profile",
    y = "Mean score"
  ) +
  theme(
    legend.position = "none",
    strip.text = element_text(size = 9)
  )

# boxplot + individual jittered points, so we see both distribution and individual variation
ggplot(sleep_long, aes(x = Class, y = Value, fill = Class)) +
  geom_boxplot(outlier.shape = NA, alpha = .7, na.rm = TRUE) +
  geom_jitter(width = .08, alpha = .25, size = 1, na.rm = TRUE) +
  facet_wrap(~ Variable, scales = "free_y") +
  theme_classic() +
  labs(
    title = "Sleep, mood, and fatigue scores by attachment profile",
    x = "Attachment profile",
    y = "Score"
  ) +
  theme(
    legend.position = "none",
    strip.text = element_text(size = 9)
  )

# manually calculating mean + 95% CI, connecting with lines to see the trend
plot_means <- sleep_long %>%
  group_by(Class, Variable) %>%
  summarise(
    mean = mean(Value, na.rm = TRUE),
    se = sd(Value, na.rm = TRUE) / sqrt(sum(!is.na(Value))),
    .groups = "drop"
  )

ggplot(plot_means, aes(x = Class, y = mean, group = 1)) +
  geom_point(size = 2) +
  geom_line() +
  geom_errorbar(aes(ymin = mean - 1.96 * se, ymax = mean + 1.96 * se), width = .15) +
  facet_wrap(~ Variable, scales = "free_y") +
  theme_classic() +
  labs(
    title = "Mean patterns across attachment profiles",
    x = "Attachment profile",
    y = "Mean score"
  ) +
  theme(
    strip.text = element_text(size = 9)
  )
```

```{r}
# a simplified version of the mean + CI plot (no color, cleaner look)
ggplot(sleep_long, aes(x = Class, y = Value)) +
  stat_summary(fun = mean, geom = "point", size = 2.5) +
  stat_summary(fun.data = mean_cl_normal, geom = "errorbar", width = .15) +
  facet_wrap(~ Variable, scales = "free_y") +
  theme_classic() +
  labs(
    title = "Mean sleep component scores by attachment profile",
    x = "Attachment profile",
    y = "Mean score"
  )
```

```{r}
# full descriptive stats (mean, sd, median, etc.) by class for each sleep/mood variable
sleep_desc <- map_dfr(sleep_vars, \(v) {
  analysis_data %>%
    group_by(Class) %>%
    summarise(
      Variable = v,
      n = sum(!is.na(.data[[v]])),
      mean = mean(.data[[v]], na.rm = TRUE),
      sd = sd(.data[[v]], na.rm = TRUE),
      median = median(.data[[v]], na.rm = TRUE),
      IQR = IQR(.data[[v]], na.rm = TRUE),
      min = min(.data[[v]], na.rm = TRUE),
      max = max(.data[[v]], na.rm = TRUE),
      .groups = "drop"
    ) %>%
    dplyr::select(Variable, everything())
})

sleep_desc %>%
  mutate(
    across(c(mean, sd, median, IQR, min, max), ~ round(.x, 2))
  )
```

```{r}
# Welch ANOVA for each sleep/mood variable, checking if classes differ,
# then correcting p-values for multiple comparisons with BH
sleep_welch <- map_dfr(sleep_vars, \(v) {
  dat <- analysis_data %>%
    dplyr::select(Class, value = all_of(v)) %>%
    filter(!is.na(value))

  wt <- oneway.test(value ~ Class, data = dat, var.equal = FALSE)

  tibble(
    Variable = v,
    N_used = nrow(dat),
    statistic = unname(wt$statistic),
    df1 = unname(wt$parameter[1]),
    df2 = unname(wt$parameter[2]),
    p = wt$p.value
  )
}) %>%
  mutate(
    p_BH = p.adjust(p, method = "BH"),
    statistic = round(statistic, 3),
    df1 = round(df1, 2),
    df2 = round(df2, 2),
    p = round(p, 3),
    p_BH = round(p_BH, 3)
  )

sleep_welch
```

```{r}
# Games-Howell post hoc for each variable (pairs well with Welch ANOVA since it
# also doesn't assume equal variances) - checking exactly which classes differ
library(rstatix)
sig_sleep_vars <- sleep_welch %>%
  # filter(p_BH < .05) %>%
  pull(Variable)

sleep_posthoc <- if (length(sig_sleep_vars) > 0) {
  map_dfr(sig_sleep_vars, \(v) {
    analysis_data %>%
      select(Class, value = all_of(v)) %>%
      filter(!is.na(value)) %>%
      games_howell_test(value ~ Class) %>%
      mutate(Variable = v, .before = 1)
  })
} else {
  tibble()
}

sleep_posthoc %>%
  mutate(
    across(where(is.numeric), ~ round(.x, 3))
  )
```

##### Tables

```{r}
# packaging all these test results into one "paper-ready" table:
# each cell is mean (SE) + a significance letter - different letters means significant difference
library(dplyr)
library(tidyr)
library(purrr)
library(rstatix)
library(multcompView)

analysis_data <- analysis_data %>%
  mutate(Class = factor(Class))

class_levels <- levels(analysis_data$Class)

sleep_desc2 <- map_dfr(sleep_vars, \(v) {
  analysis_data %>%
    group_by(Class) %>%
    summarise(
      Variable = v,
      n = sum(!is.na(.data[[v]])),
      mean = mean(.data[[v]], na.rm = TRUE),
      se = sd(.data[[v]], na.rm = TRUE) / sqrt(n),
      .groups = "drop"
    ) %>%
    select(Variable, everything())
})

sig_vars <- sleep_welch %>%
  filter(p_BH < .05) %>%
  pull(Variable)

# figuring out the significance letters for each variable:
# if the variable itself isn't significant, just mark everything "a" (all the same)
make_letters <- function(v) {

  if (!(v %in% sig_vars)) {
    return(
      tibble(
        Variable = v,
        Class = factor(class_levels, levels = class_levels),
        Letters = "a"
      )
    )
  }

  gh <- analysis_data %>%
    select(Class, value = all_of(v)) %>%
    filter(!is.na(value)) %>%
    games_howell_test(value ~ Class)

  p_col <- if ("p.adj" %in% names(gh)) "p.adj" else "p"

  pvals <- gh[[p_col]]
  names(pvals) <- paste(gh$group1, gh$group2, sep = "-")

  letters <- multcompView::multcompLetters(
    pvals,
    threshold = 0.05
  )$Letters

  tibble(
    Variable = v,
    Class = factor(names(letters), levels = class_levels),
    Letters = letters
  ) %>%
    right_join(
      tibble(Class = factor(class_levels, levels = class_levels)),
      by = "Class"
    ) %>%
    mutate(
      Variable = v,
      Letters = ifelse(is.na(Letters), "a", Letters)
    ) %>%
    select(Variable, Class, Letters)
}

sleep_letters <- map_dfr(sleep_vars, make_letters)

# putting it all together into the final wide table - rows are variables,
# columns are the four profiles
sleep_table <- sleep_desc2 %>%
  left_join(sleep_letters, by = c("Variable", "Class")) %>%
  mutate(
    Class = recode(as.character(Class),
      "1" = "Dismissive-leaning",
      "2" = "Fearful-avoidant",
      "3" = "Anxious-preoccupied",
      "4" = "Secure"
    ),
    mean = round(mean, 2),
    se = round(se, 2),
    Cell = paste0(mean, " (", se, ")", Letters)
  ) %>%
  select(Variable, Class, Cell) %>%
  pivot_wider(
    names_from = Class,
    values_from = Cell
    # names_prefix = "Class "
  )

sleep_table
```

```{r}
# same content, but making a nicer web-friendly version with gt (superscript letters)
library(gt)

sleep_table_gt <- sleep_desc2 %>%
  left_join(sleep_letters, by = c("Variable", "Class")) %>%
  mutate(
    Class = recode(as.character(Class),
      "1" = "Dismissive-leaning",
      "2" = "Fearful-avoidant",
      "3" = "Anxious-preoccupied",
      "4" = "Secure"
    ),
    mean = round(mean, 2),
    se = round(se, 2),
    Cell = paste0(mean, " (", se, ")<sup>", Letters, "</sup>")
  ) %>%
  select(Variable, Class, Cell) %>%
  pivot_wider(
    names_from = Class,
    values_from = Cell
    # names_prefix = "Class "
  ) %>%
  gt() %>%
  fmt_markdown(columns = c(
    "Dismissive-leaning",
    "Fearful-avoidant",
    "Anxious-preoccupied",
    "Secure"
  )) %>%
  tab_source_note(
    source_note = md(
      "Classes that do not share a superscript are significantly different based on Games–Howell post-hoc tests, *p* < .05."
    )
  )

sleep_table_gt
```

```{r}
# merging the descriptives and Welch ANOVA results into one summary table
sleep_summary <- sleep_desc %>%
  group_by(Variable) %>%
  summarise(
    total_nonmissing = sum(n),
    .groups = "drop"
  ) %>%
  left_join(sleep_welch, by = "Variable")

sleep_summary
```

```{r}
# double-checking missingness for these sleep variables against Class again
sleep_missing <- map_dfr(sleep_vars, \(v) {
  dat <- analysis_data %>%
    mutate(miss = is.na(.data[[v]]))

  tab <- table(dat$Class, dat$miss)
  pval <- tryCatch(chisq.test(tab)$p.value, error = function(e) NA_real_)

  tibble(
    Variable = v,
    N_used = sum(!dat$miss),
    Missing_p_by_Class = pval
  )
})

sleep_missing
```

#### Sleep Total Score

```{r}
ggplot(analysis_data, aes(x = Class, y = SLEEP_SCORE_TOTAL, fill = Class)) +
  geom_boxplot(outlier.shape = NA, alpha = .7, na.rm = TRUE) +
  geom_jitter(width = .08, alpha = .25, size = 1, na.rm = TRUE) +
  theme_classic() +
  labs(
    title = "Total sleep score by attachment profile",
    x = "Attachment profile",
    y = "SLEEP_SCORE_TOTAL"
  ) +
  theme(legend.position = "none")
```

```{r}
analysis_data %>%
  group_by(Class) %>%
  summarise(
    n = sum(!is.na(SLEEP_SCORE_TOTAL)),
    mean = round(mean(SLEEP_SCORE_TOTAL, na.rm = TRUE), 2),
    sd = round(sd(SLEEP_SCORE_TOTAL, na.rm = TRUE), 2),
    median = round(median(SLEEP_SCORE_TOTAL, na.rm = TRUE), 2),
    IQR = round(IQR(SLEEP_SCORE_TOTAL, na.rm = TRUE), 2),
    min = round(min(SLEEP_SCORE_TOTAL, na.rm = TRUE), 2),
    max = round(max(SLEEP_SCORE_TOTAL, na.rm = TRUE), 2),
    .groups = "drop"
  )
```

```{r}
# checking whether the total sleep score differs significantly across the four attachment profiles
sleep_total_dat <- analysis_data %>%
  select(Class, SLEEP_SCORE_TOTAL) %>%
  filter(!is.na(SLEEP_SCORE_TOTAL))

oneway.test(SLEEP_SCORE_TOTAL ~ Class, data = sleep_total_dat, var.equal = FALSE)
```

```{r}
# figuring out exactly which two profiles differ
library(rstatix)

sleep_total_dat %>%
  games_howell_test(SLEEP_SCORE_TOTAL ~ Class)
```

## Subquestion

### Create age groups

```{r}
# subquestion: does the attachment profile's effect on sleep change depending on age group?
# starting simple - splitting age into "18-29" and "29+" groups
library(dplyr)
library(tidyr)
library(ggplot2)
library(purrr)
library(car)
library(emmeans)
library(sandwich)
library(DescTools)
library(scales)

analysis_data <- analysis_data %>%
  mutate(
    AgeGroup = case_when(
      age >= 18 & age <= 29 ~ "Emerging adults (18–29)",
      age > 29 ~ "Older adults (>29)",
      TRUE ~ NA_character_
    ),
    AgeGroup = factor(
      AgeGroup,
      levels = c("Emerging adults (18–29)", "Older adults (>29)")
    ),
    Class = factor(Class)
  )
```

```{r}
tab_age_class <- table(analysis_data$AgeGroup, analysis_data$Class)
tab_age_class

prop.table(tab_age_class, margin = 1)   # row proportions within age group
prop.table(tab_age_class, margin = 2)   # column proportions within class
```

```{r}
ggplot(analysis_data, aes(x = AgeGroup, fill = Class)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = label_percent()) +
  theme_classic() +
  labs(
    title = "Attachment profile composition by age group",
    x = NULL,
    y = "Proportion",
    fill = "Profile"
  )
```

### Chi-square test

```{r}
# checking whether age group and attachment profile are independent (unrelated) -
# chi-square test + effect size
chi_age_class <- chisq.test(tab_age_class)
chi_age_class
DescTools::CramerV(tab_age_class)

# standardized residuals - which cell is furthest off from what we'd expect
round(chi_age_class$stdres, 2)
```

```{r}
# running a separate two-proportion test per attachment profile, checking how
# much its share differs between the two age groups
profile_prop_tests <- map_dfr(levels(analysis_data$Class), \(cl) {
  dat <- analysis_data %>%
    filter(!is.na(AgeGroup), !is.na(Class)) %>%
    mutate(in_profile = Class == cl)

  counts <- tapply(dat$in_profile, dat$AgeGroup, sum)
  totals <- table(dat$AgeGroup)

  pt <- prop.test(x = counts, n = totals)

  tibble(
    Profile = cl,
    Emerging_prop = counts[1] / totals[1],
    Older_prop = counts[2] / totals[2],
    p = pt$p.value
  )
}) %>%
  mutate(
    p_BH = p.adjust(p, method = "BH"),
    Emerging_prop = round(Emerging_prop, 3),
    Older_prop = round(Older_prop, 3),
    p = round(p, 3),
    p_BH = round(p_BH, 3)
  )

profile_prop_tests
```

```{r}
# outcome variables we're looking at next (sleep/mood related)
outcomes <- c(
  "SLEEP_DUR_SCORE",
  "ONSET_SCORE",
  "SLP_QUALITY_SCORE",
  "Sleep_Problem_score",
  "Enthusiasm_and_trouble_score",
  "sleep_efficiencySCORE",
  "SLEEP_HYG_SCORE",
  "Anxiousmood",
  "Depressedmood",
  "fatigue"
)
```

### Descriptives by age group and profile

```{r}
# descriptives for every outcome variable, grouped by "age group x attachment profile"
outcome_desc <- analysis_data %>%
  select(AgeGroup, Class, all_of(outcomes)) %>%
  pivot_longer(
    cols = all_of(outcomes),
    names_to = "Variable",
    values_to = "Value"
  ) %>%
  group_by(Variable, AgeGroup, Class) %>%
  summarise(
    n = sum(!is.na(Value)),
    mean = mean(Value, na.rm = TRUE),
    sd = sd(Value, na.rm = TRUE),
    median = median(Value, na.rm = TRUE),
    IQR = IQR(Value, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    mean = round(mean, 2),
    sd = round(sd, 2),
    median = round(median, 2),
    IQR = round(IQR, 2)
  )

outcome_desc
```

### Graph the outcomes

```{r}
outcome_long <- analysis_data %>%
  select(AgeGroup, Class, all_of(outcomes)) %>%
  pivot_longer(
    cols = all_of(outcomes),
    names_to = "Variable",
    values_to = "Value"
  )

# coloring by age group to see if the mean trend across attachment profiles
# looks consistent between the two age groups
ggplot(outcome_long, aes(x = Class, y = Value, color = AgeGroup, group = AgeGroup)) +
  stat_summary(fun = mean, geom = "point",
               position = position_dodge(width = .3), size = 2.5, na.rm = TRUE) +
  stat_summary(fun.data = mean_cl_normal, geom = "errorbar",
               position = position_dodge(width = .3), width = .15, na.rm = TRUE) +
  facet_wrap(~ Variable, scales = "free_y") +
  theme_classic() +
  labs(
    title = "Sleep, mood, and fatigue outcomes by profile and age group",
    x = "Attachment profile",
    y = "Mean score",
    color = "Age group"
  )
```

### AgeGroup × Class for each outcome

```{r}
# using sum contrasts (effect coding) so the main effects in the two-way ANOVA are easier to interpret
options(contrasts = c("contr.sum", "contr.poly"))
```

```{r}
# a function to run a two-way "age group x attachment profile" ANOVA for each outcome variable
# (using robust HC3 standard errors so it doesn't depend as much on equal-variance assumptions)
run_two_way <- function(var) {
  dat <- analysis_data %>%
    select(AgeGroup, Class, y = all_of(var)) %>%
    filter(!is.na(AgeGroup), !is.na(Class), !is.na(y))

  fit <- lm(y ~ AgeGroup * Class, data = dat)
  aov_tab <- car::Anova(fit, type = 3, white.adjust = "hc3")

  tibble(
    Variable = var,
    N_used = nrow(dat),
    AgeGroup_F = aov_tab["AgeGroup", "F value"],
    AgeGroup_p = aov_tab["AgeGroup", "Pr(>F)"],
    Class_F = aov_tab["Class", "F value"],
    Class_p = aov_tab["Class", "Pr(>F)"],
    Interaction_F = aov_tab["AgeGroup:Class", "F value"],
    Interaction_p = aov_tab["AgeGroup:Class", "Pr(>F)"]
  )
}

# running this for every outcome variable, then correcting p-values across
# them with BH
two_way_results <- map_dfr(outcomes, run_two_way) %>%
  mutate(
    AgeGroup_p_BH = p.adjust(AgeGroup_p, method = "BH"),
    Class_p_BH = p.adjust(Class_p, method = "BH"),
    Interaction_p_BH = p.adjust(Interaction_p, method = "BH")
  ) %>%
  mutate(
    across(ends_with("_F"), ~ round(.x, 3)),
    across(ends_with("_p"), ~ round(.x, 3))
  )

two_way_results
```

```{r}
# focusing specifically on Sleep_Problem_score - pairwise comparisons across
# the four attachment profiles
# (using robust HC3 standard errors, BH-corrected for multiple comparisons)
fit_sleep <- lm(Sleep_Problem_score ~ AgeGroup * Class, data = analysis_data)

emmeans(
  fit_sleep,
  ~ Class,
  vcov. = sandwich::vcovHC(fit_sleep, type = "HC3")
) %>%
  pairs(adjust = "BH")
```
